Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-19 paper/intro.tex
wordsmithing on intro

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

BriefCASE incorporates model-level cyber-analysis tools (presently
GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
examine AADL models for potential vulnerabilities and suggest
cyber-security requirements to mitigate them.  A library of
architectural transforms guides the system engineer through automated
model transformations that modify the architecture to address these
requirements, possibly inserting new high-assurance components into
the system.  Implementations for the new components are synthesized
from formal specifications using
SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
(Semantic Properties for Language and Automata Theory).  Formal
verification that the transformed system model meets its requirements
is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
AGREE is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect (and alert) unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the incorporation of these high-assurance components, the hardened system meets its cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML source code is exactly preserved in the final binaries. Assuming that the execution schedule of the deployed cyber-hardened system is as intended by the AADL model, and that the HAMR-generated communication fabric delivers messages between components as expected, the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

CASE was scoped to address the more common cyber-vulnerabilities that programmers introduce, such as overflow, lack of input validation, supply-chain issues, and identity verification; but not other cyber-vulnerabilities such as side-channel attacks and denial of service.
In this program, we have not invented a new type of filter or monitor in terms of capability, instead our contribution is in the automated synthesis of these components from a formal specification, and a means to show that the synthesis is correct.

The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}. Contract specification in the AGREE language is presented in Section~\ref{sec:agree}, followed by a description of the synthesis pathway in Section~\ref{sec:synthesis}. A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the simple example and shows the viability of the modeling approach to a full-scale industrial design.

>>>>

The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

BriefCASE incorporates model-level cyber analysis tools (presently
GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
examine AADL models for potential vulnerabilities and suggest
cyber-security requirements to mitigate them.  A library of
architectural transforms guides the system engineer through automated
model transformations that modify the architecture to address these
requirements, possibly inserting new high-assurance components into
the system.  Implementations for the new components are synthesized
from formal specifications using
SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
(Semantic Properties for Language and Automata Theory).  Formal
verification that the transformed system model meets its requirements
is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
AGREE is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

Two security-enhancing transformations are discussed in the remainder
of the paper: (1) the insertion of a filter to prevent malformed data
from a malicious actor from being propagated to downstream components,
and (2) the insertion of a monitor to detect (and alert) unexpected
behaviors arising from untrusted components. These transformations not
only change the architecture of the model by adding in new components;
they also generate a formal specification for the behavior of the
inserted high-assurance components in the AGREE language. Those
specifications are sufficient for model checking to prove that---due
to the newly included high-assurance components---the hardened
system meets its cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE
specifications for the high-assurance components to CakeML, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from specifications to CakeML code, providing a formal framework in
which to argue correctness. CakeML then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the CakeML source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the AGREE model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

CASE was scoped to address common cyber-vulnerabilities, such as
overflow, lack of input validation, supply-chain issues, and identity
verification; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are yet dealt with in our
work.  Research in CASE has not invented a new type of filter or
monitor, in terms of capability; instead, our contribution is in the
automated synthesis of these components from a formal specification,
and a means to show that the synthesis is correct.

The approach is illustrated by a simple example in
Section~\ref{sec:example}. Contract specification in the AGREE
language is presented in Section~\ref{sec:agree}, followed by a
description of the synthesis pathway in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.
